### 2018-06-30 paper/1-intro.tex
Tweak S1 and S2

In this paper we offer a general framework in which to understand
and compare build systems,
in a way that is both abstract (omitting incidental detail)
and yet precise (implemented as Haskell code).  Specifically we make
these contributions:
\begin{itemize}
\item Build systems vary on many axes, including:
static vs dynamic dependencies; local vs cloud; deterministic vs
non-deterministic build rules; support for early cutoff; self-tracking build
systems; and the type of persistent build information.
In \S\ref{sec-background} we identify some key properties, in the context of
four carefully-chosen build systems.
\item We describe some simple but novel abstractions that
  crisply encapsulate what a build system is (\S\ref{sec-abstractions}),
  allowing us, for example, to speak about what it means for a build system to be correct.

\item We identify two key design choices
  that are typically deeply wired into any build system:
  \emph{the order in which dependencies are built} (\S\ref{sec-dependency-orderings})
  and \emph{whether or not a dependency is (re-)built} (\S\ref{sec-out-of-date}).
  These choices turn out to be orthogonal, which leads us to a new
  classification of the design space (\S\ref{sec-design-space}).

>>>>

In this paper we offer a general framework in which to understand
and compare build systems,
in a way that is both abstract (omitting incidental detail)
and yet precise (implemented as Haskell code:
\url{https://hackage.haskell.org/package/build-0.0.1.1}). Specifically we make
these contributions:
\begin{itemize}
\item Build systems vary on many axes, including: static vs dynamic
  dependencies; local vs cloud; deterministic vs non-deterministic build rules;
  support for early cutoff; self-tracking build systems; and the type of
  persistent build information. In~\S\ref{sec-background} we identify some key
  properties, in the context of four carefully-chosen build systems.

\item We describe some simple but novel abstractions that
  crisply encapsulate what a build system is (\S\ref{sec-abstractions}),
  allowing us, for example, to speak about what it means for a build system to be correct.

\item We identify two key design choices that are typically deeply wired into
  any build system: \emph{the order in which dependencies are
  built}~(\S\ref{sec-dependency-orderings}) and \emph{whether or not a
  dependency is (re-)built}~(\S\ref{sec-out-of-date}). These choices turn out to
  be orthogonal, which leads us to a new classification of the design
  space~(\S\ref{sec-design-space}).

### 2018-06-30 paper/2-background.tex
Tweak S1 and S2

If the user runs \Make specifying \cmd{main.exe} as the desired output, \Make
will first build \cmd{util.o} and \cmd{main.o}, in any order since these
tasks are independent, and then build \cmd{main.exe}. If the
user modifies the sources of \cmd{util.h} and runs \Make again, it will
perform a \emph{full rebuild}, because all three tasks transitively depend on
\cmd{util.h}, as illustrated in Fig.~\ref{fig-make}(b). On the other hand, if
the user modifies \cmd{main.c} then a \emph{partial rebuild} is sufficient:
the file \cmd{util.o} does not need to be rebuilt, since its inputs have not
changed, see Fig.~\ref{fig-make}(c). Note that if the dependency graph is
\emph{acyclic} then each task needs to be executed at most once. Cyclic task
dependencies are typically not allowed in build systems although there are rare
exceptions, see~\S\ref{sec-iterative-compute}.

The following property is essential for build systems; indeed, it is their raison d'\^etre:

@@

\begin{figure}[h]
\vspace{-2mm}
\begin{subfigure}[b]{0.90\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/shake-example.pdf}}
\caption{Dependency graph produced after the previous build.}
\end{subfigure}
\begin{subfigure}[b]{0.90\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/shake-example-rebuild.pdf}}
\caption{Since \cmd{docs.txt} was modified, we rebuild \cmd{release.txt} and
\cmd{release.tar}, discovering a new dependency.}
\end{subfigure}
\vspace{-2mm}
\caption{Dynamic dependencies example: create \cmd{README} and add it to the
list of release documents \cmd{docs.txt}.\label{fig-shake}}
\vspace{-4mm}
\end{figure}

@@

\Shake's implementation is different from both \Make and \Excel in two aspects.
First, it uses the dependency graph from the previous build to decide which
files need to be rebuilt. This idea has a long history, going back to
\emph{incremental}~\cite{demers1981incremental},
\emph{adaptive}~\cite{acar2002adaptive}, and
\emph{self-adjusting computations} (see \cite{acar2007selfadjusting} and
\S\ref{sec-related}). Second, instead of aborting and deferring the execution of
tasks whose newly discovered dependencies have not yet been built (as \Excel
does), \Shake \emph{suspends} their execution until the dependencies are brought
up to date. We refer to this task scheduling algorithm as \emph{suspending}.

\begin{figure}[h]
\vspace{-5mm}
\centerline{\includegraphics[scale=0.28]{fig/shake-example-cutoff.pdf}}
\vspace{-3mm}
\caption{An early cutoff example: if a comment is added to \cmd{main.c}, the
rebuild is stopped after detecting that \cmd{main.o} is unchanged, since this
indicates that \cmd{main.exe} and its dependents do not need to be
rebuilt.\label{fig-cutoff}}
\vspace{-2mm}
\end{figure}

@@

\begin{figure}[t]
\begin{subfigure}[b]{0.25\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/bazel-example-checkout.pdf}}
\caption{Download source files}
\end{subfigure}
\begin{subfigure}[b]{0.40\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/bazel-example-build.pdf}}
\caption{Build \cmd{main.exe}}
\end{subfigure}
\begin{subfigure}[b]{0.31\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/bazel-example-rebuild.pdf}}
\caption{Modify \cmd{util.c} and rebuild}
\end{subfigure}
\caption{A cloud build example: (a)~download sources, (b)~build \cmd{main.exe}
by downloading it from the cloud and skipping intermediate files (only their
hashes are needed), (c)~modify \cmd{util.c} and rebuild \cmd{main.exe}, which
requires building \cmd{util.o} (nobody has compiled \cmd{util.c} before) and
downloading \cmd{main.o} (it is needed for linking \cmd{main.exe}). File hashes
are shown in circles, and non-materialised intermediates in dashed rounded
rectangles.
\label{fig-bazel}}
\end{figure}

@@

\begin{table}[h]
\smaller
\centering
\begin{tabular}{l||l|l||l|c|c|c}
\hline
$\!$Build system$\!$& Persistent build information & Scheduler   & Dependencies    & Minimal & Cutoff & Cloud$\!$\\\hline
$\!$\Make       $\!$& File modification times      & Topological & Static          & Yes     & No     & No   $\!$\\
$\!$\Excel      $\!$& Dirty cells, calc chain      & Restarting  & Dynamic         & No      & No     & No   $\!$\\
$\!$\Shake      $\!$& Previous dependency graph    & Suspending  & Dynamic         & Yes     & Yes    & No   $\!$\\
$\!$\Bazel      $\!$& Cloud cache, command history & Restarting  & Dynamic$^{(*)}$ & No      & Yes    & Yes  $\!$\\\hline
\hline
\end{tabular}
\vspace{0.5mm}
\caption{Summary of build system differences; $^{(*)}$user-defined build rules
cannot have dynamic dependencies.\label{tab-summary}}
\vspace{-6mm}
\end{table}

@@

>>>>

If the user runs \Make specifying \cmd{main.exe} as the desired output, \Make
will build~\cmd{util.o}~and~\cmd{main.o}, in any order since these tasks are
independent, and then \cmd{main.exe}. If the user modifies \cmd{util.h} and runs
\Make again, it will perform a \emph{full rebuild}, because all three tasks
transitively depend on \cmd{util.h}, as illustrated in Fig.~\ref{fig-make}(b).
On the other hand, if the user modifies \cmd{main.c} then a \emph{partial
rebuild} is sufficient: \cmd{util.o} does not need to be rebuilt, since its
inputs have not changed, see Fig.~\ref{fig-make}(c). Note that if the dependency
graph is \emph{acyclic} then each task needs to be executed at most once. Cyclic
task dependencies are typically not allowed in build systems but there are rare
exceptions, see~\S\ref{sec-iterative-compute}.

\noindent
The following property is essential for build systems; indeed, it is their
raison d'\^etre:

@@

\begin{figure}
\begin{subfigure}[b]{0.90\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/shake-example.pdf}}
\caption{Dependency graph produced after the previous build.}
\end{subfigure}
\begin{subfigure}[b]{0.90\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/shake-example-rebuild.pdf}}
\caption{Since \cmd{docs.txt} was modified, we rebuild \cmd{release.txt} and
\cmd{release.tar}, discovering a new dependency.}
\end{subfigure}
\vspace{-2mm}
\caption{Dynamic dependencies example: create \cmd{README} and add it to the
list of release documents \cmd{docs.txt}.\label{fig-shake}}
\vspace{-4mm}
\end{figure}

@@

\Shake's implementation is different from both \Make and \Excel in two aspects.
First, it uses the dependency graph from the previous build to decide which
files need to be rebuilt. This idea has a long history, going back to
\emph{incremental}~\cite{demers1981incremental},
\emph{adaptive}~\cite{acar2002adaptive}, and
\emph{self-adjusting computations} (see \cite{acar2007selfadjusting} and
\S\ref{sec-related}). Second, instead of aborting and deferring the execution of
tasks whose newly discovered dependencies have not yet been built (as \Excel
does), \Shake \emph{suspends} their execution until the dependencies are brought
up to date. We refer to this task scheduling algorithm as \emph{suspending},
see a concrete implementation in~\S\ref{sec-implementation-shake}.

\begin{figure}
\vspace{-2mm}
\centerline{\includegraphics[scale=0.28]{fig/shake-example-cutoff.pdf}}
\vspace{-3mm}
\caption{An early cutoff example: if a comment is added to \cmd{main.c}, the
rebuild is stopped after detecting that \cmd{main.o} is unchanged, since this
indicates that \cmd{main.exe} and its dependents do not need to be
rebuilt.\label{fig-cutoff}}
\vspace{-4mm}
\end{figure}

@@

\begin{figure}
\begin{subfigure}[b]{0.25\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/bazel-example-checkout.pdf}}
\caption{Download source files}
\end{subfigure}
\begin{subfigure}[b]{0.40\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/bazel-example-build.pdf}}
\caption{Build \cmd{main.exe}}
\end{subfigure}
\begin{subfigure}[b]{0.31\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/bazel-example-rebuild.pdf}}
\caption{Modify \cmd{util.c} and rebuild}
\end{subfigure}
\vspace{-2mm}
\caption{A cloud build example: (a)~download sources, (b)~build \cmd{main.exe}
by downloading it from the cloud and skipping intermediate files (only their
hashes are needed), (c)~modify \cmd{util.c} and rebuild \cmd{main.exe}, which
requires building \cmd{util.o} (nobody has compiled \cmd{util.c} before) and
downloading \cmd{main.o} (it is needed for linking \cmd{main.exe}). File hashes
are shown in circles, and non-materialised intermediates in dashed rounded
rectangles.\label{fig-bazel}}
\vspace{-4mm}
\end{figure}

@@

@@

\begin{table}
\smaller
\centering
\begin{tabular}{l||l|l||l|c|c|c}
\hline
$\!$Build system$\!$& Persistent build information & Scheduler   & Dependencies    & Minimal & Cutoff & Cloud$\!$\\\hline
$\!$\Make       $\!$& File modification times      & Topological & Static          & Yes     & No     & No   $\!$\\
$\!$\Excel      $\!$& Dirty cells, calc chain      & Restarting  & Dynamic         & No      & No     & No   $\!$\\
$\!$\Shake      $\!$& Previous dependency graph    & Suspending  & Dynamic         & Yes     & Yes    & No   $\!$\\
$\!$\Bazel      $\!$& Cloud cache, command history & Restarting  & Dynamic$^{(*)}$ & No      & Yes    & Yes  $\!$\\\hline
\hline
\end{tabular}
\vspace{0.5mm}
\caption{Build system differences; $^{(*)}$at present, user-defined build rules
cannot have dynamic dependencies.\label{tab-summary}}
\vspace{-6mm}
\end{table}
